Algebraic Topology Final Report: Morse Theory

### 2013-01-12 paper.tex
doing sec 6

Let $M$ be a smooth manifold of dimension $k$. By Whitney Embedding Theorem, we
can embed $M$ into $\RR^n, n > k$. Let $N = \{ (q,v) \in \RR^{2n} | q \in M, v
\in (T_qM)^{\perp} \subset \RR^n \}$ be the normal bundle of $M$ in $\RR^n$. We
define a map $E$ that maps $N$ into the ambient space $\RR^n$ by
\begin{equation}
  E(q, v) = q + v.
\end{equation}
The point $q+v$ thus line in a normal line that is perpendicular to $M$ at
$p$. This leads to the definition of focal points:

\begin{defn}
  e \in \RR^n is a focal point of (M, q) with multiplicity \mu 
\end{defn}

\section{Manifolds in Euclidean Space: The Existence of Non-degenerate
  Functions}

>>>>

\section{Manifolds in Euclidean Space: The Existence of Non-degenerate
  Functions}
Let $M$ be a smooth manifold of dimension $k$. By Whitney Embedding Theorem, we
can embed $M$ into $\RR^n, n > k$. Therefore, without loss of generality we may
assume $M \subset \RR^n$. For each $p \in \RR^n$, define a smooth function $L_p
: M \to \RR $ by
\begin{equation}
  L_p(q) = |q-p|^2.
\end{equation}

We want to find a $p \in \RR^n$ such that $L_p$ has no degenerate critical
points in $M$. Let $q \in M$ and let $(u^1, .., u^k) \mapsto (x^1(u), ..,
x^n(u))$ be a local coordinate at $q$. Then $q$ is a critical point if
$\frac{\partial f}{\partial u^i}(q) = 0 \forall i = 1, .., k.$; that is,

Let $N = \{ (q,v) \in \RR^{2n} | q \in M, v \in (T_qM)^{\perp} \subset \RR^n \}$
be the normal bundle of $M$ in $\RR^n$. Then $N$ is a manifold of dimension
$k+(n-k)=n$. We define a map $E$ that maps $N$ into the ambient space $\RR^n$ by
\begin{equation}
  E(q, v) = q + v.
\end{equation}
This is a differentiable map between two manifolds of dimension $n$. The point
$q+v$ thus line in a normal line that is perpendicular to $M$ at $p$. This leads
to the definition of focal points:

\begin{defn}
  $e \in \RR^n$ is a focal point of $(M, q)$ with multiplicity $\mu > 0$ if $e =
  q+v$, where $(q, v) \in N$ and the Jacobian of $E$ at $(q, v)$ has nullity
  $\mu > 0$. $e$ is called a focal point of $M$ if $e$ is a focal point of $(M,
  q)$, $q \in M$.
\end{defn}

The focal points will play an important role in the construction of a kind
